---
title: "Assignment4"
author: "Madeline Gorchels"
date: "11/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, message=FALSE}
#Loading packages and data sets (see note below)

library(tidyverse)
library(effsize)
library(knitr)
library(kableExtra)
library(vcdExtra) # Based on instructions it looks like this may be necissary -you'll need to instal this if you haven't already

lob_ab = read_csv("lobster_size_abundance.csv")
lob_trap = read_csv("lobster_traps.csv")

```

# "Lobster abundance and fishing pressure (2012 - 2017)"

Describe trends in lobster abundance (counts) and fishing pressure (trap buoys) at the five locations from
2012 - 2017. Ignore transect information - we are only interested in evaluating abundance and pressure on the
order of SITE. Note: you are not expected to use regression here - just think of ways to clearly
describe annual totals visually and in text, noting important trends, events and differences.


```{r}

tidy_lob_ab <- expand.dft(data.frame(lob_ab), freq = "COUNT") # Data in tidy format

lob_ab_hist <- ggplot(tidy_lob_ab, aes(SIZE)) +
  geom_histogram(aes(fill = SITE)) +
  facet_wrap(~SITE)

lob_ab_hist

lob_ab_hist_stack <- ggplot(tidy_lob_ab, aes(SIZE)) +
  geom_histogram(aes(fill = SITE)) 

lob_ab_hist_stack
```

```{r}
lob_trap_filtered <- filter(lob_trap, SITE == c("AQUE","CARP","IVEE","MOHK","NAPL"))  # I get a warning message so I need to confirm this is ok

lob_trap_hist <- ggplot(lob_trap_filtered, aes(TRAPS)) +
  geom_histogram(aes(fill = SITE)) +
  facet_wrap(~SITE)

lob_trap_hist

```



# "Compare mean lobster size by site in 2017"

```{r}
lob_size_2017 <- select(tidy_lob_ab, YEAR, SITE, SIZE) %>%  # Select relevant categories. Note, depending on how the we make the file tidy, the first term may need to change 
  filter(YEAR == "2017") %>% 
  group_by(SITE)
```

Test to see if the pushing works for me (Becky)



